Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants with Transient Performance Specifications

### 2017-06-15 automatica/paper.tex

\section{Introduction}
\textcolor{red}{Cristina: We have to write a new introduction as well.}

%\begin{figure}
%\begin{center}
%\includegraphics[height=4cm]{jcaesar.eps}    % The printed column  
%\caption{Gaius Julius Caesar, 100--44 B.C.}  % width is 8.4 cm.
%\label{fig1}                                 % Size the figures 
%\end{center}                                 % accordingly.
%\end{figure}

% OR

%\begin{figure}
%\begin{center}
%\epsfig{file=jcaesar,width=7cm}
%\caption{Gaius Julius Caesar, 100--44 B.C.}
%\label{fig1}
%\end{center}
%\end{figure}

%\subsection{A subsection}
%Marcus Tullius Cicero, 106--43 B.C. was a Roman statesman, orator, 
%and philosopher.  A major figure in the last years of the Republic, 
%he is best known for his orations against Catiline\footnote{
%This footnote should be very brief.}
%and for his mastery of Latin prose \cite{Heritage:92}. He was a 
%contemporary of Julius Caesar (Fig.~\ref{fig1}).

>>>>

\section{Introduction}
Modern implementations of embedded control systems have proliferated with
the availability of low-cost devices that can perform highly non-trivial
control tasks, with significant impact in numerous application areas such as
life sciences, environmental control and robotics~\cite{astrom1997computer, Franklin15}. 
Correct synthesis is non-trivial, even in cases with restriceted dynamics.
We examine the case of Linear Time Invariant (LTI) Models, for which the
synthesis of controllers is well understood. However, the
use of digital control architectures adds new challenges caused by artifacts
specific to digital control, such as the effects of finite-precision arithmetic, time
discretization and quantization noise introduced by A/D and D/A conversion.

While research on digital control is well developed~\cite{astrom1997computer},
automated and sound control synthesis is challenging, particularly when the
synthesis objective goes beyond classical stability.  There are recent methods
for verifying reachability properties of a given controller~\cite{FLD+11}.
However, these methods have not been generalized to control synthesis.  
Note that a synthesis algorithm that guarantees stability does not
ensure safety: the system might transitively visit an unsafe state resulting
in unrecoverable failure. Programming expertise is a key
barrier to broad adoption of correct digital controllers, and requires
considerable knowledge outside of the expertise of many control engineers.

Beyond classical a-posteriori validation in digital control, there has been
plenty of previous work aiming at \emph{verifying} a given designed
controller, which however broadly lack automation.  Recent work has studied
the stability of digital controllers considering implementation aspects,
i.e., fixed-point arithmetic and the word length~\cite{Bessa16}.  They
exploit advances in bit-accurate verification of C programs to obtain a
verifier for software-implemented digital control.

By contrast, we leverage a very recent step-change in the automation and
scalability of \emph{program synthesis}.  Program synthesis engines use a
specification as the starting point, and subsequently generate a sequence of
candidate programs from a given template.  The candidate programs are
iteratively refined to eventually satisfy the specification.  Program
synthesizers implementing Counter-Example Guided Inductive Synthesis
(CEGIS)~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06} are now able to generate
programs for highly non-trivial specifications with a very high degree of
automation.  Modern synthesis engines combine automated testing, genetic
algorithms, and SMT-based automated
reasoning~\cite{DBLP:journals/corr/AlurFSS16a, DBLP:conf/lpar/DavidKL15}.

By combining state-of-the-art synthesis engines we present a tool that
automatically generates digital controllers for a given continuous plant
model that are correct by construction.  This approach delivers a high
degree of automation, promises to reduce the cost and time of development of
digital control dramatically, and requires considerably less expertise than
a-posteriori verification.  Specifically, we synthesize stable,
software-implemented embedded controllers along with a model of a physical
plant.  Due to the complexity of such closed-loop systems, in this work we
focus on linear models with known configurations, and perform parametric
synthesis of stabilizing digital controllers with a number of properties that
define their desired behaviour.

Our work addresses challenging aspects of the control synthesis problem.
We perform digital control synthesis over a hybrid model, where the plant
exhibits continuous behavior whereas the controller operates in discrete
time and over a quantized domain. In particular, our model evaluates the
effects of the quantizers (A/D and D/A converters) and of time discretization,
as well as representation errors introduced by in modelling artifacts
such as an observer~\cite{astrom1997computer} in a finite precision domain, 
while reasoning about high level properties such as stability, robustness, safety,
and response time, which define the behaviour of the system. In particular,
safety requirements are frequently overlooked in conventional feedback control
synthesis, but play an important role in systems engineering.

Using state of the art verifiers, we evaluate these models against automatically 
synthesised controllers using CEGIS-based synsthesis that incorporate the above
mentioned artifacts in different models of Single-Input/Single-Output (SISO) systems.
We offer three modes of operation that address different aspects of the problem.

{\em The first  approach} uses a transfer-function based analysis that solves the
problem of robust stability for closed loop systems for a collection of unstable plants.
Inthis case, there is no safety specification since we are working in the frequency domain.

{\em The second approach} uses a na\"ive technique that starts by devising a
digital controller that stabilizes the system while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded safety by unfolding the dynamics of the system, considering the
full hyper-cube of initial states, and checking a {\em completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
system such that the boundaries are not violated for any larger number of
iterations.  As~it requires unfolding up to the completeness threshold, this
approach is computationally expensive.

Instead of unfolding the dynamics, {\em the third approach}
employs {\em abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the system simultaneously. 
Additionally, this approach uses {\em abstraction refinement},
enabling us to always start with a very simple description regardless of the
dynamics complexity, and only expand to more complex models
when a solution cannot be found.

We provide experimental results showing that all our approaches are able to
efficiently synthesize safe controllers for a set of intricate
physical plant models taken from the digital control literature.

In summary, this paper makes the following original contributions.
%
\begin{enumerate}

\item We automatically generate {\em correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  In particular, we
  have found that a two-stage verification engine that continuously refines
  the precision of the fixed-point representation of the plant yields a
  speed-up of two orders of magnitude over a conventional one-stage
  verification engine.

%
\item  We compute state-feedback controllers that 
  guarantee a given safety property alongside stability and response time
  specifications.  Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide three novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  A third algorithm
  evaluates systems using transfer functions where no state-space description is given.
  All three approaches provide sound synthesis of feedback controllers and consider
  the various sources of imprecision in the implementation of the control
  algorithm and in the modeling of the plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.
\end{enumerate}
